---
title: "01 - Flowlines"
author: "Angus Watters"
date: "2023-05-13"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
library(ggmapinset)
library(sf)
library(ggplot2)
library(dplyr)
library(sfheaders)
library(hydrofab)

knitr::opts_chunk$set(echo = TRUE)
options(warn = -1, verbose = FALSE)
source("../nhdplusv2/utils.R")

# getting VAA from NHDPlus
vaa <- nhdplusTools::get_vaa()

# new attributes
new_atts = read_parquet("D:/nhdplusv2_reference_features/enhd_nhdplusatts.parquet")
# new_atts = read_parquet(glue("{base_dir}/enhd_nhdplusatts.parquet"))

# Get the current flowline path
fl_path <- "D:/nhdplusv2_reference_features/02_Flowlines/NHDPlus06.gpkg"

ble_path <- "D:/nhdplusv2_reference_features/02_BLE/NHDPlus06.gpkg"

# Get the VPU identifier for the current flowline file
# which_VPU = gsub(".gpkg", "", gsub("NHDPlus", "", basename(fl_path)))

# # Get the BLE path for the current VPU
# ble_path  = ble_paths[grep(which_VPU, basename(ble_paths))]
# 
# # Define the output file path for the processed flowlines
# outfile   = glue("{reference_dir}flowlines_{which_VPU}.gpkg")

# inset map point
lng = c(-84.17691)
lat = c(35.15163)

pt <-   data.frame(lng = lng, 
                   lat = lat
                   ) %>% 
  sf::st_as_sf( coords = c("lng", "lat"), crs = 4326) %>% 
  sf::st_geometry() %>% 
  sf::st_centroid()

# ggplot theme 
thm <- 
  ggplot2::theme_bw() +
  ggplot2::theme(
    plot.title = ggplot2::element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle =  ggplot2::element_text(size = 14, hjust = 0.5)
  )

group_colors <- c(
  "19485688" = "#00BA38",
  "19485662" = "#F8766D",
  "19485570" = "#619CFF"
  )
```

The goal of this step in generating a new and improved NHDPlus dataset is to rectify missing attributes in NHDPlus flowlines and resolve network connectivity issues. 

To show what we did, we will walk through an example of our process on VPU 6. 

<br>
<br>

# 1. Read in NHDPlus flowlines

We first read in our original NHDPlus flowlines dataset, drop any Z/M dimensions, and align our columns with the NHDPlus names.

```{r, eval = T, echo = T}
# read in flowlines
nhd <- sf::read_sf(fl_path)

# remove Z/M dimensions and align names
nhd <- 
  nhd %>% 
  sf::st_zm() %>%
  nhdplusTools::align_nhdplus_names()
```

<br>
<br>

# 2. Join with VAA and E2NHD 

We have 2 sets of attribute data that we want to join and with `nhd`:

1. Value Added Attribute set. 
2. Updated network attributes from E2NHD https://doi.org/10.5066/P9W79I7Q

Here is a list of the attributes in the the Value Added Attributes dataset
```{r}
# getting VAA from NHDPlus
vaa <- nhdplusTools::get_vaa()

names(vaa)
```

And these are the attributes in the E2NHD.
```{r}
# new attributes
new_atts <- arrow::read_parquet("D:/nhdplusv2_reference_features/enhd_nhdplusatts.parquet")

names(new_atts)
```

We first join `vaa` with `nhd` to attach network connectivty information to each of the flowlines COMIDs (i.e. from/to nodes, start flag, stream calc #)

```{r, eval = TRUE, echo = T}
nhd <-
  nhd %>% 
  dplyr::left_join(
    vaa, 
    by = c("COMID" = "comid")
    ) %>%
  dplyr::select(COMID, fromnode, tonode, startflag, streamcalc, divergence, dnminorhyd)
```

<br>

We then join the `nhd` dataset with the `new_atts` from E2NHD which contain improved network connectivity information. For more information visit https://doi.org/10.5066/P9W79I7Q. We also realign the names to match the NHDPlus names, and recalculate the flowline lengths. 

```{r, eval = TRUE, echo = T}
nhd <-
  nhd %>% 
  dplyr::left_join(
    new_atts,
    by = c("COMID" = "comid")
    ) %>%
  nhdplusTools::align_nhdplus_names() %>%
  dplyr::mutate(LENGTHKM  = add_lengthkm(.))
```

Now we have a NHDPlus flowlines with the necessary attributes.

<br>
<br>

# 3. Burn Lines

Next thing to do is get our Burn lines data (`ble`), and rename the `LineID` column to `COMID`.

```{r, eval = TRUE, echo = T}
# Read the BLE file and perform some preprocessing steps
ble <- sf::read_sf(ble_path) 

# rename LineID column to COMID
ble <- dplyr::rename(ble, COMID = LineID)

ble
```

<br>
<br>

# 4. Join Burn Lines data with `nhd data`
We then join our the burn line multilinestring geometries with the `nhd` flowlines COMIDs. We also drop any Z/M dimensions. Once we have done this, we are left with `ble` data that uses the burn line multilinestring as the geometries for any matching COMIDs, all other geometries will be represented by empty geometries.

<br>

In our example, all of the geometries are empty (i.e. There are no COMIDs in `ble` that match COMIDs in `nhd`)
```{r}
# Join the NHDPlus flowlines with the BLE lines where applicable
ble <- 
  dplyr::left_join(
    dplyr::select(sf::st_drop_geometry(nhd), COMID), 
    ble, 
    by = "COMID"
  ) %>%
  sf::st_as_sf() %>%
  sf::st_zm()
```

<br>

 We then create a `flag` for non empty geometries, where the `nhd` data has a `startFlag` equal to 1 or a `Divergence` equal to 2. This flag indicates where to use the burn line geometries.
 
```{r, eval = TRUE, echo = TRUE}
# Define a flag to indicate where the BLE lines will be used
flag <- !sf::st_is_empty(sf::st_geometry(ble)) & (nhd$StartFlag == 1 | nhd$Divergence == 2)
```

```{r, eval = TRUE, echo = TRUE}
table(flag)
```

In our example, all flowlines where flagged as FALSE for these conditions.

<br>
<br>

# 5. Replace any flagged geometries with burn line geometries
We then replace any geometries in `nhd` with the burn line multilinestrings that were flagged as non empty start flags or divergences.  
```{r, eval = TRUE, echo = TRUE}
# Replace the flowline geometry with the BLE line geometry where applicable
sf::st_geometry(nhd)[flag] <- sf::st_geometry(ble)[flag]
```

<br>
<br>

# 6. Create a custom network
Using our `nhd` data, we use our `FromNode` and `ToNode` for each COMID to get a corrected `tocomid` that we will use to override the current `nhd` `toCOMID` data.
```{r, eval = TRUE, echo = TRUE}
# Generate a custom network for the flowlines based on the override_tocomid values
custom_net <- 
  nhd %>% 
  sf::st_drop_geometry() %>%
  dplyr::select(COMID, FromNode, ToNode, Divergence) %>%
  nhdplusTools::get_tocomid(remove_coastal = FALSE) %>%
  dplyr::select(comid, override_tocomid = tocomid)
```

<br>

We then join this `custom_net` with our `nhd` data by COMID to add the `override_tocomid` column to our original data. We then update the `override_tocomid` data by setting the value to the `override_tocomid` from the `custom_net` if the `toCOMID` value equals 0, otherwise we just keep the original `toCOMID` COMID value.
```{r, eval = TRUE, echo = TRUE}
# Update the override_tocomid values to account for nodes with no downstream flow
nhd <- 
  dplyr::left_join(
    nhd, 
    custom_net, 
    by = c("COMID" = "comid")
  ) %>%
  dplyr::mutate(
    override_tocomid = ifelse(toCOMID == 0, override_tocomid, toCOMID)
  )
```


```{r, eval = TRUE, echo = TRUE}
# Check for any invalid override_tocomid values, that each flowline is a headwater and leads somewhere
check <- !nhd$COMID %in% nhd$override_tocomid &
  !(nhd$override_tocomid == 0 | is.na(nhd$override_tocomid) |
      !nhd$override_tocomid %in% nhd$COMID)

# Filter the flowlines to only those with
check_direction <- filter(nhd, check)
```


<br>
<br>

















